---
title: "DataCleaning-add some change"
output: word_document
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

------------------------------------------------------------------------

# Library initialization

```{r libraries, message=FALSE}
library(tidyverse)
library(readr)
library(VIM)
library(DMwR2)
library(Hmisc)
library(PTXQC)
library(stringr)
```

# Compiled cleanup

### compile all clean-up

#### re-read data

```{r copy-data}
clean = read_csv(file = 'vehicles.csv')
cor(clean$odometer,clean$price,use = 'complete.obs')
```

#### remove unused column

```{r unused-column}
clean <- select(clean,c(-VIN,-url,-image_url,-region_url,-county))

```

#### remove dirty rows

**first remove rows contain NAs above 20%**

```{r drop rows with many NAs}
clean <- clean[-manyNAs(clean,nORp = 0.2),]
```

```{r remove-nonsense-row}
clean = clean[!clean$model %in% c("0","50鈥檚"),]
clean = clean[!is.na(clean$year),]
clean = clean[!is.na(clean$price),]
```

```{r remove-outliers}
clean = clean[clean$year >= 1998 & clean$year <= 2020,] 
clean = clean[clean$price > 1988 & clean$price < 44999,] 
```

#### format data

```{r format-data}
clean$region[clean$region == "columbia / jeff city"] <- "columbia"
clean$region[clean$region == "corvallis/albany"] <- "albany"
clean$region[clean$region == "dayton / springfield"] <- "springfield"
clean$region[clean$region == "florence / muscle shoals"] <- "florence"
clean$region[clean$region == "huntsville / decatur"] <- "decatur"
clean$region[clean$region == "lafayette / west lafayette"] <- "lafayette"

clean$model[clean$model == "124 spider $25k"] <- "124 spider" 
clean$model[clean$model == "z4 3.0i 3.0i"] <- "z4 3.0i"
clean$model[clean$model == "ZL1 Camaro"] <- "zl1 camaro" 
clean$model[clean$model == "cx9 touring"] <-"cx-9 touring"
clean$model[clean$model == "cmax"] <-"c-max"
clean$model[clean$model == "cmax hybrid"] <-"c-max hybrid"
clean$model[clean$model == "camaro iroc-z28"] <-"camaro iroc z28"
clean$model[clean$model == "fx35 suv"] <-"fx35"
clean$model[clean$model == "PLYMOUTH BELVEDERE"] <-"plymouth belvedere"
clean$model[clean$model == "cx9 grand touring"] <-"cx-9 grand touring"
clean$model[clean$model == "cx9"] <-"cx-9"
clean$model[clean$model == "cx7"] <-"cx-7"
clean$model[clean$model == "cls550"] <-"cls 550"
clean$model[clean$model == "s-class s 550"] <-"s-class s550"
clean$model[clean$model == "escape se eco boost"] <-"sescape se ecoboost"
clean$model[clean$model == "devville"] <-"deville"
clean$model[clean$model == "explorer xls"] <-"explorer xlt"
clean$model[clean$model == "EXPLORER"] <-"explorer"
clean$model[clean$model == "e150"] <-"e-150"
clean$model[clean$model == "e150 conversion van"] <-"e-150 conversion van"
clean$model[clean$model == "e350 super duty"] <-"e-350 super duty"
clean$model[clean$model == "f350 super duty"] <-"f-350 super duty"
clean$model[clean$model == "f350 super duty xlt"] <-"f-350 super duty xlt"
clean$model[clean$model == "f-350 xl"] <-"f350 xl"
clean$model[clean$model == "genesis 3.8l 4dr sedan"] <-"genesis 3.8 sedan 4d"
clean$model[clean$model == "imiev"] <-"i-miev"
clean$model[clean$model == "impeza"] <-"impreza"
clean$model[clean$model == "riveira"] <-"riviera"
clean$model[clean$model == "rs7"] <-"rs 7"
clean$model[clean$model == "is300"] <-"300"
clean$model[clean$model == "jetta 2.5l se"] <-"jetta 2.5 se"
clean$model[clean$model == "INTERNATIONAL 4700"] <-"international 4700"
clean$model[clean$model %in% c("PETERBUILT 379", "peterbilt 379")] <-"Peterbilt 379"    
clean$model[clean$model %in% c("pilot exl", "pilot ex-l | 4wd, 3rd row")] <-"pilot ex-l"   
clean$model[clean$model %in% c("Cars,Trucks,Vans,Suvs","cars suvs trucks", "Cars/trucks/suvs")] <-"cars-trucks-vans-suvs" 
clean$model[clean$model %in% c("CONVERTIBLE", "convert", "convertable")] <-"convertible"      
clean$model[clean$model %in% c("international 4300", "INTERNATIONAL 4300")] <-"International 4300"     
clean$model[clean$model %in% c("IS 250", "is250")] <-"is 250"   
clean$model[clean$model %in% c("john deer", "John Deere")] <-"john deere"       
clean$model[clean$model %in% c("rav 4", "rav-4")] <-"rav4"          
clean$model[clean$model %in% c("E350", "e350")] <-"e-350"        
clean$model[clean$model %in% c("srt-10", "srt 10")] <-"srt10"
```

```{r impute-columns}
clean$title_status <- as.character(impute(clean$title_status,'missing')) 
clean$manufacturer <- as.character(impute(clean$manufacturer,'unknown'))  
clean$model <- as.character(impute(clean$model,'unknown'))    
clean$condition <- as.character(impute(clean$condition,'unknown'))
clean$cylinders <- as.character(impute(clean$cylinders,'unknown'))    
clean$description <- impute(clean$description,'none')
clean$fuel <- as.character(impute(clean$fuel,'unknown'))    
clean$transmission <- as.character(impute(clean$transmission,'unknown'))     
```

***add a new column[posting_Year],refine the method of dealing with odometer***

```{r clean-odometer}
clean$posting_date <- format(as.Date(clean$posting_date, format="%d/%m/%Y"),
                      "%d/%m/%Y")
clean['posting_year'] <- as.numeric(format(as.Date(clean$posting_date,
                                        format="%d/%m/%Y"),"%Y"))
clean['posting_month'] <- as.numeric(format(as.Date(clean$posting_date,
                                        format="%d/%m/%Y"),"%m"))
df_odo <- clean[clean$odometer < 7209 | clean$odometer > 204142|is.na(clean$odometer),]
df_odo$year_diff <- df_odo$posting_year-df_odo$year
df_odo$odometer <- (df_odo$posting_year-df_odo$year)*12000+
  df_odo$posting_month*1000
clean[clean$odometer < 7209 | clean$odometer > 204142 | is.na(clean$odometer),]$odometer = df_odo$odometer
clean <- select(clean, -posting_month) 
```

for abnormal values and NA, change their odometer to age of cars multiply by 12,000 (which is considered as average mileage per year of a vehicle)

------------------------------------------------------------------------

## ***Addition 1.0***

#### Check remained NAs

```{r clean check na}
colSums(is.na(clean))
```

#### Drop rows: lat/long contain small NAs, just drop dirty rows.

```{r drop lat/long missing rows}
clean <- clean[!is.na(clean$lat),]
clean <- clean[!is.na(clean$long),]
```

#### Impute: drive/size/type/paint_color with 'unknown'

```{r impute drive/size/type/paint_color}
clean$drive <- impute(clean$drive,'unknown')
clean$size <- impute(clean$size,'unknown')
clean$type <- impute(clean$type,'unknown')
clean$type[clean$type == "other"] <-"unknown"
clean$paint_color <- impute(clean$paint_color,'unknown')
```

#### Count Nas --0

```{r}
sum(is.na(clean))
```

## ***Addition 2.0***

### extract unknown data from description

-   for manufacturer, condition, cylinders, fuel, title_status, transmission, drive, size, type, paint_color.

-   this part will **waste huge time**, can just download the cleaned dataset from git.

1.  manufacturer

```{r extract manufacturer}
df.unknown = clean[clean$manufacturer=='unknown',]
li_descr=as.list(df.unknown$description)
make <- unique(clean$manufacturer)
make <- make[make!='unknown']
for (i in li_descr){
  for (j in make){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$manufacturer <- j
    }
  }
}
clean[clean$manufacturer=='unknown',] <- df.unknown
```

2.  condition

```{r extract condition}
df.unknown = clean[clean$condition=='unknown',]
li_descr = as.list(df.unknown$description)
condition <- unique(clean$condition)
condition <- condition[condition!='unknown']
for (i in li_descr){
  for (j in condition){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$condition <- j
    }
  }
}
clean[clean$condition=='unknown',]$condition <- df.unknown$condition
```

3.  cylinders

```{r extract cylinders}
df.unknown = clean[clean$cylinders=='unknown',]
li_descr = as.list(df.unknown$description)
cylinders <- unique(clean$cylinders)
cylinders <- cylinders[cylinders!='unknown' & cylinders!='other']
for (i in li_descr){
  for (j in cylinders){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$cylinders <- j
    }
  }
}
clean[clean$cylinders=='unknown',]$cylinders <- df.unknown$cylinders
```

4.  fuel

```{r extract fuel}
df.unknown = clean[clean$fuel=='unknown',]
li_descr = as.list(df.unknown$description)
fuel <- unique(clean$fuel)
fuel <- fuel[fuel!='unknown' & fuel!='other']
for (i in li_descr){
  for (j in fuel){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$fuel <- j
    }
  }
}
clean[clean$fuel=='unknown',]$fuel <- df.unknown$fuel
```

5.  title_status

```{r extract title_status}
df.unknown = clean[clean$title_status=='missing',]
li_descr = as.list(df.unknown$description)
title_status <- unique(clean$title_status)
title_status <- title_status[title_status!='missing']
for (i in li_descr){
  for (j in title_status){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$title_status <- j
    }
  }
}
clean[clean$title_status=='missing',]$title_status <- df.unknown$title_status
clean['title_status'][clean['title_status'] == 'missing'] <- 'unknown'
```

6.  transmission

```{r extract transmission}
df.unknown = clean[clean$transmission=='unknown',]
li_descr = as.list(df.unknown$description)
transmission <- unique(clean$transmission)
transmission <- transmission[transmission!='unknown' & transmission!='other']
for (i in li_descr){
  for (j in transmission){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$transmission <- j
    }
  }
}
clean[clean$transmission=='unknown',]$transmission <- df.unknown$transmission
```

7.  drive

```{r extract drive}
df.unknown = clean[clean$drive=='unknown',]
li_descr = as.list(df.unknown$description)
drive <- unique(clean$drive)
drive <- drive[drive!='unknown']
for (i in li_descr){
  for (j in drive){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$drive <- j
    }
  }
}
clean[clean$drive=='unknown',]$drive <- df.unknown$drive
```

8.  size

```{r extract size}
df.unknown = clean[clean$size=='unknown',]
li_descr = as.list(df.unknown$description)
size <- unique(clean$size)
size <- size[size!='unknown']
for (i in li_descr){
  for (j in size){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$size <- j
    }
  }
}
clean[clean$size=='unknown',]$size <- df.unknown$size
```

9.  type

```{r extract type}
df.unknown = clean[clean$type=='unknown',]
li_descr = as.list(df.unknown$description)
type <- unique(clean$type)
type <- type[type!='unknown' & type!='other']
for (i in li_descr){
  for (j in type){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$type <- j
    }
  }
}
clean[clean$type=='unknown',]$type <- df.unknown$type
```

10. paint_color

```{r extract paint_color}
df.unknown = clean[clean$paint_color=='unknown',]
li_descr = as.list(df.unknown$description)
paint_color <- unique(clean$paint_color)
paint_color <- paint_color[paint_color!='unknown']
for (i in li_descr){
  for (j in paint_color){
    if (str_detect(i,j)==T){
      df.unknown[df.unknown$description==i,]$paint_color <- j
    }
  }
}
clean[clean$paint_color=='unknown',] <- df.unknown
```

#### drop description column

```{r drop descr}
clean = select(clean,-description)
```

#### add make_model column

```{r}
clean['make_model'] <- paste(clean$manufacturer,clean$model,sep=':')
```

#### Output clean data

```{r}
write.csv(clean, "clean-vehicles.csv", row.names = F)
write.csv(clean, "./shiny_vehicles/heatmap/clean-vehicles.csv", row.names = F)
```

## ***Misc.***

#### Read data

```{r read data}
data=read.csv("clean-vehicles.csv", stringsAsFactors = T)
# use stringsAsFactors when reading csv
```

#### See the structure of data

```{r}
glimpse(data)
```
