---
title: "fuel-VIN cleaning"
output: word_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## My part contains 5 variables' cleaning decisions, devided into 6 parts.

**Descriptions** (may be useful when creating the codebook):

**fuel** - Type of fuel used in vehicles.

**odometer** - The distance traveled by a vehicle.

**title_status** - a legal form refers to the ownership of a vehicle.

**transmission** - Automatic or manual transmission.

**VIN** - Vehicle Identification Number.

## Part 0: A general look at the dataset.

read in libraries

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(VIM)
library(DMwR2)
library(Hmisc)
```

read in our data

\*Notice: Use read_csv() instead of read.csv() because some columns contain "," in original csv file.

```{r}
df <- read_csv(file = 'vehicles.csv') 
```

```{r, echo=FALSE}
print(paste('total missing values:',sum(is.na(df))))
```

#### See the distribution of missing values.

```{r, VIM}
aggr(df,labels=names(df),col=c('blue','red'),
     numbrs=T,sortVars=T,Cex.axis=.7,gap=3)
```
Can see 'county' column is a missing column and 71.7% values in 'size' are missing.



## Part 1: clean fuel

#### compute missing values of fuel

```{r, echo=FALSE}
print(paste('NAs of fuel:',sum(is.na(df$fuel))))
```

#### Plan: Just drop all rows which fuel is missing.

```{r clean fuel}
d_c_fuel <- df[complete.cases(df$fuel),]
sum(is.na(d_c_fuel$fuel))
table(d_c_fuel$fuel)
```



## Part 2: clean VIN

#### compute missing values of fuel

```{r, echo=FALSE}
print(paste('NAs of VIN',sum(is.na(d_c_fuel$VIN))))
```

#### Plan: Drop the 'VIN' column because it contains large missing values, and each obs. has a unique value, which is difficult to extract insights. 

```{r clean VIN}
d_c_VIN <- select(d_c_fuel,-VIN)
dim(d_c_VIN)
```



## Part 3: clean transmission

#### compute missing values of transmission

```{r, echo=FALSE}
print(paste('NAs of transmission',
            sum(is.na(d_c_VIN$transmission))))
```

#### Plan: Just drop all rows which transmission is missing.

```{r clean transmission}
d_c_transmission <- d_c_VIN[complete.cases(d_c_VIN$transmission),]
sum(is.na(d_c_transmission$transmission))
table(d_c_transmission$transmission)
```



## Part 4: clean title_status

#### compute missing values of title_status

```{r, echo=FALSE}
print(paste('NAs of title_status',
            sum(is.na(d_c_transmission$title_status))))
```

see factor levels of title_status

```{r}
table(d_c_transmission$title_status)
```
/*contain 505 obs. with 'missing' label.

#### Plan: Impute NAs of title_status with 'missing'.

```{r clean title_status}
d_c_title_status <- d_c_transmission
d_c_title_status$title_status <- as.character(impute(d_c_title_status$title_status,'missing'))
sum(is.na(d_c_title_status$title_status))
table(d_c_title_status$title_status)
```



## Part 5: clean odometer

compute missing values of odometer

```{r, echo=FALSE}
print(paste('NAs of odometer:',
            sum(is.na(d_c_title_status$odometer))))
```
#### Plan: deal with outliers and impute NAs.

**use boxplot to see the distribution of outliers**

```{r}
odo_bp <- boxplot(d_c_title_status$odometer,boxwex=0.7)
odo_bp$stats
```

**impute NAs by mean**

```{r}
d_c_odometer <- d_c_title_status
d_c_odometer[is.na(d_c_odometer$odometer),]$odometer <- odo_bp$stats[3]
```

**smooth outliers**

```{r}
d_c_odometer[d_c_odometer$odometer>odo_bp$stats[5],]$odometer <- odo_bp$stats[5]
max(d_c_odometer$odometer)
boxplot(d_c_odometer$odometer)
```










