---
title: "Data Cleaning"
output: word_document
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library initialization

```{r libraries, message=FALSE}
library(tidyverse)
library(readr)
library(VIM)
library(DMwR2)
library(Hmisc)
library(PTXQC)
```

# General look at the Dataset

#### Read file

\*Notice: Use read_csv() instead of read.csv() because some columns contain "," in original csv file.

```{r file-read}
data <- read_csv(file = 'vehicles.csv')  
```

#### Total missing values

```{r total-missing, echo=FALSE}
sum(is.na(data))
```

#### Missing values distribution

```{r na-distribution, VIM}
aggr(data,labels=names(data),col=c('blue','red'),
     numbrs=T,sortVars=T,Cex.axis=.7,gap=3)
```

# By Column Cleaning

## Clean ID column

#### ID Column contain neither duplicate or n/a

```{r clean-id}
nrow(data[duplicated(data$id),]) 
nrow(data[is.na(data$id),]) 
```

## Clean url

#### url column contain no n/a or duplicate

```{r clean-url}
nrow(data[duplicated(data$url),]) 
nrow(data[is.na(data$url),]) 
```

## Clean Region

#### region column contain no n/a

```{r clean-region}
nrow(data[is.na(data$region),])
```

### analyze region with '/'

#### replace (region1 / region2) =\> (region1) for redundant regions

```{r replace-region}
a_region = unique(data[!grepl('/', data$region),]$region,)
x_region = unique(data[grepl('/', data$region),]$region,) 

xa_region = a_region[grepl(paste(unlist(lapply(strsplit(x_region, "/"), str_trim)), collapse="|"), a_region)]  
f_region = x_region[grepl(paste(c(a_region, xa_region), collapse="|"), x_region)]  

d_c_region = data
d_c_region$region[d_c_region$region == "columbia / jeff city"] <- "columbia"
d_c_region$region[d_c_region$region == "corvallis/albany"] <- "albany"
d_c_region$region[d_c_region$region == "dayton / springfield"] <- "springfield"
d_c_region$region[d_c_region$region == "florence / muscle shoals"] <- "florence"
d_c_region$region[d_c_region$region == "huntsville / decatur"] <- "decatur"
d_c_region$region[d_c_region$region == "lafayette / west lafayette"] <- "lafayette"

unique(d_c_region$region)[order(unique(d_c_region$region))]
```

## Clean region_url column

#### region_url column contain no n/a

```{r clean-url}
nrow(data[is.na(data$region_url),]) 
```

## Clean Price column

#### remove missing price (na / 0) and display removed row count from original data

```{r clean-price}
d_c_price = data[!is.na(data$price) & data$price != 0,]
nrow(data[data$price == 0,]) 
```

## Clean Fuel column

**fuel** - Type of fuel used in vehicles.
#### Missing values

```{r missing-fuel, echo=FALSE}
sum(is.na(data$fuel))
```

#### Drop all rows with missing fuel

```{r clean-fuel}
d_c_fuel <- data[complete.cases(data$fuel),]
sum(is.na(d_c_fuel$fuel))
table(d_c_fuel$fuel)
```

## Clean VIN

**VIN** - Vehicle Identification Number.
#### compute missing values of fuel

```{r missing-VIN, echo=FALSE}
sum(is.na(d_c_fuel$VIN))
```

#### Drop the 'VIN' column because it contains large missing values, and each obs. has a unique value, which is difficult to extract insights.

```{r clean VIN}
d_c_VIN <- select(d_c_fuel,-VIN)
dim(d_c_VIN)
```

## Clean Transmission

**transmission** - Automatic or manual transmission.
#### missing values of transmission

```{r missing-transmission, echo=FALSE}
sum(is.na(d_c_VIN$transmission))
```

#### drop all rows with missing transmission.

```{r clean-transmission}
d_c_transmission <- d_c_VIN[complete.cases(d_c_VIN$transmission),]
sum(is.na(d_c_transmission$transmission))
table(d_c_transmission$transmission)
```

## Clean title_status column

**title_status** - a legal form refers to the ownership of a vehicle.
#### missing values of title_status

```{r missing-title-status, echo=FALSE}
sum(is.na(d_c_transmission$title_status))
```

#### factor levels of title_status

```{r factor-levels-title-status}
table(d_c_transmission$title_status)
```

/\*contain 505 obs.
with 'missing' label.
#### Impute NAs of title_status with 'missing'.

```{r clean title_status}
d_c_title_status <- d_c_transmission
d_c_title_status$title_status <- as.character(impute(d_c_title_status$title_status,'missing'))
sum(is.na(d_c_title_status$title_status))
table(d_c_title_status$title_status)
```

## Clean odometer column

**odometer** - The distance traveled by a vehicle.
#### missing values of odometer

```{r missing-odometer, echo=FALSE}
print(paste('NAs of odometer:',
            sum(is.na(d_c_title_status$odometer))))
```

### deal with outliers and impute NAs.

#### use boxplot to see the distribution of outliers

```{r odometer-boxplot}
odo_bp <- boxplot(d_c_title_status$odometer,boxwex=0.7)
odo_bp$stats
```

#### impute NAs by mean

```{r odometer-impute}
d_c_odometer <- d_c_title_status
d_c_odometer[is.na(d_c_odometer$odometer),]$odometer <- odo_bp$stats[3]
```

#### smooth outliers

```{r odometer-smooth}
d_c_odometer[d_c_odometer$odometer>odo_bp$stats[5],]$odometer <- odo_bp$stats[5]
max(d_c_odometer$odometer)
boxplot(d_c_odometer$odometer)
```

## Clean image_url column

#### remove attribute 'image_url' entirely

## Clean drive column

#### all drive column data is shown acceptable

```{r unique-drives}
unique(data$drive)
```

## Clean size column

#### all size column data is shown acceptable

```{r unique-size}
unique(data$size)
```

## Clean type column

#### all type column data is shown acceptable

```{r unique-type}
unique(data$type)
```

## Clean paint_color column

#### all paint_color column data is shown acceptable

```{r unique-paint-color}
unique(data$paint_color)
```

## Clean year column

#### remove missing year

```{r clean-year}
nrow(data[is.na(data$year),])  
```

#### all year column data is shown acceptable

```{r unique-year}
unique(data$year)
```

## Clean manufacturer column

#### remove missing manufacturer

```{r clean-manufacturer}
nrow(data[is.na(data$manufacturer),])  
```

#### all manufacturer column data is shown acceptable

```{r unique-manufacturer}
unique(data$manufacturer)
```

## Clean model column

#### remove missing model

```{r clean-model}
nrow(data[is.na(data$model),])  
nrow(data[data$model %in% c("0","50鈥檚"),])  

```

#### replace inconsistent value

```{r inconsistent-model}
d_c_model = data
d_c_model$model[d_c_model$model == "124 spider $25k"] <- "124 spider" 
d_c_model$model[d_c_model$model == "z4 3.0i 3.0i"] <- "z4 3.0i"
d_c_model$model[d_c_model$model == "ZL1 Camaro"] <- "zl1 camaro" 
d_c_model$model[d_c_model$model == "cx9 touring"] <-"cx-9 touring"
d_c_model$model[d_c_model$model == "cmax"] <-"c-max"
d_c_model$model[d_c_model$model == "cmax hybrid"] <-"c-max hybrid"
d_c_model$model[d_c_model$model == "camaro iroc-z28"] <-"camaro iroc z28"
d_c_model$model[d_c_model$model == "fx35 suv"] <-"fx35"
d_c_model$model[d_c_model$model == "PLYMOUTH BELVEDERE"] <-"plymouth belvedere"
d_c_model$model[d_c_model$model == "cx9 grand touring"] <-"cx-9 grand touring"
d_c_model$model[d_c_model$model == "cx9"] <-"cx-9"
d_c_model$model[d_c_model$model == "cx7"] <-"cx-7"
d_c_model$model[d_c_model$model == "cls550"] <-"cls 550"
d_c_model$model[d_c_model$model == "s-class s 550"] <-"s-class s550"
d_c_model$model[d_c_model$model == "escape se eco boost"] <-"sescape se ecoboost"
d_c_model$model[d_c_model$model == "devville"] <-"deville"
d_c_model$model[d_c_model$model == "explorer xls"] <-"explorer xlt"
d_c_model$model[d_c_model$model == "EXPLORER"] <-"explorer"
d_c_model$model[d_c_model$model == "e150"] <-"e-150"
d_c_model$model[d_c_model$model == "e150 conversion van"] <-"e-150 conversion van"
d_c_model$model[d_c_model$model == "e350 super duty"] <-"e-350 super duty"
d_c_model$model[d_c_model$model == "f350 super duty"] <-"f-350 super duty"
d_c_model$model[d_c_model$model == "f350 super duty xlt"] <-"f-350 super duty xlt"
d_c_model$model[d_c_model$model == "f-350 xl"] <-"f350 xl"
d_c_model$model[d_c_model$model == "genesis 3.8l 4dr sedan"] <-"genesis 3.8 sedan 4d"
d_c_model$model[d_c_model$model == "imiev"] <-"i-miev"
d_c_model$model[d_c_model$model == "impeza"] <-"impreza"
d_c_model$model[d_c_model$model == "riveira"] <-"riviera"
d_c_model$model[d_c_model$model == "rs7"] <-"rs 7"
d_c_model$model[d_c_model$model == "is300"] <-"300"
d_c_model$model[d_c_model$model == "jetta 2.5l se"] <-"jetta 2.5 se"
d_c_model$model[d_c_model$model == "INTERNATIONAL 4700"] <-"international 4700"
d_c_model$model[d_c_model$model %in% c("PETERBUILT 379", "peterbilt 379")] <-"Peterbilt 379"    
d_c_model$model[d_c_model$model %in% c("pilot exl", "pilot ex-l | 4wd, 3rd row")] <-"pilot ex-l"   
d_c_model$model[d_c_model$model %in% c("Cars,Trucks,Vans,Suvs","cars suvs trucks", "Cars/trucks/suvs")] <-"cars-trucks-vans-suvs" 
d_c_model$model[d_c_model$model %in% c("CONVERTIBLE", "convert", "convertable")] <-"convertible"      
d_c_model$model[d_c_model$model %in% c("international 4300", "INTERNATIONAL 4300")] <-"International 4300"     
d_c_model$model[d_c_model$model %in% c("IS 250", "is250")] <-"is 250"   
d_c_model$model[d_c_model$model %in% c("john deer", "John Deere")] <-"john deere"       
d_c_model$model[d_c_model$model %in% c("rav 4", "rav-4")] <-"rav4"          
d_c_model$model[d_c_model$model %in% c("E350", "e350")] <-"e-350"        
d_c_model$model[d_c_model$model %in% c("srt-10", "srt 10")] <-"srt10"  
```

## Clean condition column

#### remove missing condition

```{r clean-condition}
nrow(data[is.na(data$condition),])
```

#### all condition column data is shown acceptable

```{r unique-condition}
unique(data$condition)
```

## Clean cylinders column

#### remove missing cylinders

```{r clean-cylinders}
nrow(data[is.na(data$cylinders),])
```

#### all cylinders column data is shown acceptable

```{r unique-cylinders}
unique(data$cylinders)
```

# Compiled cleanup

### compile all clean-up

#### re-read data

```{r copy-data}
clean = read_csv(file = 'vehicles.csv')  
```

#### remove unused column

```{r unused-column}
clean <- select(clean,-VIN)
```

#### remove dirty rows

```{r remove-dirty-row}
clean = clean[!clean$model %in% c("0","50鈥檚"),]
clean = clean[!is.na(clean$year),]
clean = clean[!is.na(clean$price),]
clean = clean[clean$price != 0,]
clean = clean[!duplicated(clean$url),]
clean <- clean[complete.cases(clean$fuel),]
clean <- clean[complete.cases(clean$transmission),]
```

#### format data

```{r format-data}
clean$region[clean$region == "columbia / jeff city"] <- "columbia"
clean$region[clean$region == "corvallis/albany"] <- "albany"
clean$region[clean$region == "dayton / springfield"] <- "springfield"
clean$region[clean$region == "florence / muscle shoals"] <- "florence"
clean$region[clean$region == "huntsville / decatur"] <- "decatur"
clean$region[clean$region == "lafayette / west lafayette"] <- "lafayette"

clean$model[clean$model == "124 spider $25k"] <- "124 spider" 
clean$model[clean$model == "z4 3.0i 3.0i"] <- "z4 3.0i"
clean$model[clean$model == "ZL1 Camaro"] <- "zl1 camaro" 
clean$model[clean$model == "cx9 touring"] <-"cx-9 touring"
clean$model[clean$model == "cmax"] <-"c-max"
clean$model[clean$model == "cmax hybrid"] <-"c-max hybrid"
clean$model[clean$model == "camaro iroc-z28"] <-"camaro iroc z28"
clean$model[clean$model == "fx35 suv"] <-"fx35"
clean$model[clean$model == "PLYMOUTH BELVEDERE"] <-"plymouth belvedere"
clean$model[clean$model == "cx9 grand touring"] <-"cx-9 grand touring"
clean$model[clean$model == "cx9"] <-"cx-9"
clean$model[clean$model == "cx7"] <-"cx-7"
clean$model[clean$model == "cls550"] <-"cls 550"
clean$model[clean$model == "s-class s 550"] <-"s-class s550"
clean$model[clean$model == "escape se eco boost"] <-"sescape se ecoboost"
clean$model[clean$model == "devville"] <-"deville"
clean$model[clean$model == "explorer xls"] <-"explorer xlt"
clean$model[clean$model == "EXPLORER"] <-"explorer"
clean$model[clean$model == "e150"] <-"e-150"
clean$model[clean$model == "e150 conversion van"] <-"e-150 conversion van"
clean$model[clean$model == "e350 super duty"] <-"e-350 super duty"
clean$model[clean$model == "f350 super duty"] <-"f-350 super duty"
clean$model[clean$model == "f350 super duty xlt"] <-"f-350 super duty xlt"
clean$model[clean$model == "f-350 xl"] <-"f350 xl"
clean$model[clean$model == "genesis 3.8l 4dr sedan"] <-"genesis 3.8 sedan 4d"
clean$model[clean$model == "imiev"] <-"i-miev"
clean$model[clean$model == "impeza"] <-"impreza"
clean$model[clean$model == "riveira"] <-"riviera"
clean$model[clean$model == "rs7"] <-"rs 7"
clean$model[clean$model == "is300"] <-"300"
clean$model[clean$model == "jetta 2.5l se"] <-"jetta 2.5 se"
clean$model[clean$model == "INTERNATIONAL 4700"] <-"international 4700"
clean$model[clean$model %in% c("PETERBUILT 379", "peterbilt 379")] <-"Peterbilt 379"    
clean$model[clean$model %in% c("pilot exl", "pilot ex-l | 4wd, 3rd row")] <-"pilot ex-l"   
clean$model[clean$model %in% c("Cars,Trucks,Vans,Suvs","cars suvs trucks", "Cars/trucks/suvs")] <-"cars-trucks-vans-suvs" 
clean$model[clean$model %in% c("CONVERTIBLE", "convert", "convertable")] <-"convertible"      
clean$model[clean$model %in% c("international 4300", "INTERNATIONAL 4300")] <-"International 4300"     
clean$model[clean$model %in% c("IS 250", "is250")] <-"is 250"   
clean$model[clean$model %in% c("john deer", "John Deere")] <-"john deere"       
clean$model[clean$model %in% c("rav 4", "rav-4")] <-"rav4"          
clean$model[clean$model %in% c("E350", "e350")] <-"e-350"        
clean$model[clean$model %in% c("srt-10", "srt 10")] <-"srt10"
 
clean$title_status <- as.character(impute(clean$title_status,'unknown')) 
clean$manufacturer <- as.character(impute(clean$manufacturer,'unknown'))  
clean$model <- as.character(impute(clean$model,'unknown'))    
clean$condition <- as.character(impute(clean$condition,'unknown'))
clean$cylinders <- as.character(impute(clean$cylinders,'unknown'))    

odo_bp <- boxplot.stats(clean$odometer)
clean[is.na(clean$odometer),]$odometer <- odo_bp$stats[3]
clean[clean$odometer > odo_bp$stats[5],]$odometer <- odo_bp$stats[5] 
```

#### output clean data

```{r clean-output}
View(clean)
write.csv(clean, "clean_vehicles.csv")
clean
```

# 
